---
title: Apparent preferences for cognitive effort fade when multiple forms of effort and delay are interleaved in a foraging environment
author: "Claudio Toro-Serey^1^2 & Joseph T. McGuire^1^2"
csl: apa.csl
output:
  word_document: default #reference_docx: Style_reference_draft.docx
  pdf_document:
    highlight: haddock
    keep_tex: no
header-includes:
- \usepackage{setspace}\doublespacing
- \usepackage[left]{lineno}
- \linenumbers
- \usepackage{float}
- \newcommand{\beginsupplement}{ \setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
  \setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}} }
#subtitle: 'Abbreviated Title: Heterogeneity in the functional topography of mPFC'
#bibliography: mPFC_paper.bib
---

```{r Global options, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, fig.pos = 'H')
```

```{r Setup, include = FALSE}
## libraries
library(tidyverse)
library(ggfortify)
library(knitr)
library(pander) 
library(nloptr)
library(pwr)
library(lme4)
#library(lmerTest)
library(gridExtra)
library(reshape2)
library(corrplot)

## aesthetic options
lbls <- c("Wait","Cognitive","Physical","Easy") # between subj
colsBtw = c("#78AB05","#D9541A","deepskyblue4", "darkgoldenrod2") # plot colors (wait, effort)
colsWth <- c("#D9541A", "#78AB05", "dodgerblue4", "deepskyblue3")#"grey30", "grey70") # plot colors (wait, effort)
lthick = 2 # line thickness for plots

## functions
# get the number of continuous quits before current trial
seqQuits <- function(subjData = 1){
  
    # get the lag
    tempSeq <- sequence(rle(subjData)$lengths) # sequence
    tempSeq[subjData==1] <- 0 # not interested in completed runs, just quits
    tempSeq <- c(0,tempSeq) # lag it so it can be used as a t-1 regressor

    # return data so that the length is equivalent
    out <- tempSeq[1:length(subjData)]
    
    return(out)
    
}

# generic permutation for 2 groups
permute <- function(group1 = 1, group2 = 2, statType = mean, nPerms = 5000, paired = FALSE, simple = FALSE){
  
    # prep data
    summaryPerm <- list()
    lOne <- length(group1)
    lTwo <- length(group2)
    bigSample <- c(group1,group2)  
    
    if (paired == FALSE) {
      
        
        for (i in 1:nPerms){
          
            # relabel samples
            tempBig <- sample(bigSample)
            tempOne <- tempBig[1:lOne]
            tempTwo <- tempBig[(lOne+1):length(bigSample)]
            
            # stats
            tempDiffs <- statType(tempOne,na.rm=T) - statType(tempTwo,na.rm=T)
            summaryPerm$jointDist[i] <- tempDiffs # statType(tempDiffs, na.rm = T) 
          
        }  
        
    } else {
      
        for (i in 1:nPerms){
          
            # shift labels in a pairwise fashion
            tempDiffs <- statType((-1)^rbinom(lOne,1,0.5) * (group1 - group2))
            summaryPerm$jointDist[i] <- tempDiffs
          
        }
      
    }
  
    # get the observed difference
    diffs <- statType(group1,na.rm=T) - statType(group2,na.rm=T)
    observedAbs <- abs(diffs) # maybe leave it as means here
    observed <- diffs
    summaryPerm$Pval <- 2 * (1 - ecdf(summaryPerm$jointDist)(observedAbs))
    if (length(unique(abs(summaryPerm$jointDist))) == 1) {summaryPerm$Pval <- 1} # if the difference is always the same, then p = 1
    summaryPerm$Observed <- observed
    
    # return the results
    if (simple) {
      
      return(summaryPerm$Pval)
      
    } else if (!simple) {
      
      return(summaryPerm)
      
    }
    
}

# Non-parametric Bootstrap for a single group
bootstrap <- function(group = 1, statType = mean, B = 5000){
  
    # prep param
    bootStats <- rep(0,B)
    
    # iterate
    for(b in 1:B){
      
      # wait group
      x <- sample(group,length(group),replace=T)  
      bootStats[b] <- statType(x,na.rm = T)
      
    }
    
    return(bootStats)
    
}

# basic power calculation for a one way ANOVA
powerCalc <- function(d = 1.2, Za = 1.96, Zb = 0.842){
  
  out <- 2*((Za + Zb)/d)^2 + (0.25*(Za^2))
  
  return(round(out))
  
}

# OC-specific computation of the negative log likelihood for model optimization
negLogLik <- function(params, choice, handling, reward) {
  gamma <- exp(params[2]) * handling
  model <- exp(params[1]) * (reward - gamma)
  p = 1 / (1 + exp(-model))
  p[p == 1] <- 0.999
  p[p == 0] <- 0.001
  tempChoice <- rep(NA, length(choice))
  tempChoice[choice == 1] <- log(p[choice == 1])
  tempChoice[choice == 0] <- log(1 - p[choice == 0]) # log of probability of choice 1 when choice 0 occurred
  negLL <- -sum(tempChoice) 
  return(negLL)
}

# optimize the OC model
optimizeOCModel <- function(Data, Algorithm = "NLOPT_LN_NEWUOA", simplify = F, optfun = negLogLik) {
  
  # Data: The participant's log
  # Algorithm: probably let be
  # optfun: an external function to minimize (in this case OC, separately defined as negloglik)
  
  # Prep data
  handling <- Data$Handling
  reward <- Data$Offer
  choice <- Data$Choice
  
  # Prep list of results to be returned
  out <- list()
  out$percentQuit <- mean(choice == 0) * 100
  out$percentAccept <- mean(choice== 1) * 100 
  miss <- (choice != 1) & (choice != 0)
  out$percentMiss <- mean(miss)  * 100
  choice <- choice[!miss]
  reward <- as.numeric(reward[!miss])
  handling <- as.numeric(handling[!miss])
  
  # Establish lower and upper bounds
  LB <- round(log((min(reward)/max(handling)) * 0.99), digits = 4)
  UB <- round(log((max(reward)/min(handling)) * 1.01), digits = 4) # in reality this should be the second largest, since no one would reject the highest val
  
  # Begin defining parameters
  # If choices are one-sided (i.e. all accepted)
  if ((sum(choice) == length(choice)) | (sum(choice) == 0)) {
    ifelse(sum(choice) == length(choice), out$Gamma <- exp(LB), out$Gamma <- exp(UB)) 
    out$Scale <- 1 # it was NA, but in theory a temperature of 1 also indicates noiseless estimates, and allows for easier fit computations
    out$LL <- 0
  } else {
    # Create a feasible region (search space)
    params <- as.matrix(expand.grid(scale = c(-1, 1), gamma = seq(LB, UB, length = 3)))
    # Create a list to check the minimization of the negative log lik.
    info <- list()
    info$negLL <- Inf
    # Define the options to be used during optimization
    # Consider looking into other optimization algorithms and global minima
    opts <- list("algorithm" = Algorithm,
               "xtol_rel" = 1.0e-8)
    # Optimize the sOC over all possible combinations of starting points
    for (i in seq(nrow(params))) {
      tempInfo <- nloptr(x0 = params[i, ], 
                   eval_f = optfun, 
                   lb = c(log(0.001), LB),
                   ub = c(-log(0.001), UB),
                   opts = opts,
                   choice = choice, 
                   handling = handling,
                   reward = reward)
      if (tempInfo$objective < info$negLL) {
        #print("Minimized")
        info$negLL <- tempInfo$objective
        info$params <- tempInfo$solution
      }
    }
    out$Gamma <- exp(info$params[2])
    out$Scale <- exp(info$params[1])
    out$LL <- -info$negLL
  }
  
  # Summarize the outputs
  out$LL0 <- log(0.5) * length(choice)
  out$Rsquared <- 1 - (out$LL/out$LL0)
  out$subjOC <- out$Gamma * handling
  out$p <- 1 / (1 + exp(-(out$Scale*(reward - out$subjOC))))
  out$predicted <- reward > out$subjOC
  out$predicted[out$predicted == TRUE] <- 1
  out$percentPredicted <- mean(out$predicted == choice) 
  
  # adjust the probabilities in case of extreme gammas
  if (out$Gamma <= exp(LB)) {
    out$Gamma <- exp(LB) # temporary condition because Nlopt is not respecting the lower bound
    out$p <- rep(1, length(choice))
  } else if (out$Gamma == exp(UB)) {
    out$p <- rep(0, length(choice))
  } 
  
  # if doing this with dplyr::do(), return a simplified data.frame instead with the important parameters
  if (simplify) {
    out <- data.frame(out[c(seq(8), 12)])
  }   
  
  return(out)

}

# summary matrices for refrence-changing models
betaMatrix <- function(model, rearrange = NA) {
# get a similarity matrix of the resulting coefficient pairings for the cost conditions
# first, do a full_join based on column names on the list of coefficient vectors from each dummy code relevel
# then match the names of columns and rows so NAs are in the diagonal
  
  # get the names of the reference group per model iteration
  refnames <- names(model)
  
  # coefficient matrix
  temp <- lapply(model, function(data) {coefficients(data)$SubjID[1, 2:4]})
  mixCoeffs <- bind_rows(temp) 
  preln <- ifelse("Cost" %in% substr(names(mixCoeffs), 1, 4), 4, 5) # count how many characters precede the name of each cost (diff across studies)
  dimnames(mixCoeffs) <- list(refnames, substr(names(mixCoeffs), preln + 1, 20))
  mixCoeffs <- as.matrix(mixCoeffs[, match(rownames(mixCoeffs), colnames(mixCoeffs))])
  mixCoeffs[is.na(mixCoeffs)] <- 0
  
  # now the pvals
  temp <- lapply(model, function(data) {as.list(summary(data)$coefficients[2:4, 4])})
  mixPvals <- as.matrix(bind_rows(temp)) 
  dimnames(mixPvals) <- list(refnames, substr(colnames(mixPvals), preln + 1, 20))
  mixPvals <- as.matrix(mixPvals[, match(rownames(mixPvals), colnames(mixPvals))])
  mixPvals[is.na(mixPvals)] <- 1

  # if you would like to re-arrange the coefficient order, supply a vector with the desired sequence
  if (length(rearrange) > 1) {
    mixCoeffs <- mixCoeffs[rearrange, rearrange]
    dimnames(mixCoeffs) <- list(rearrange, rearrange)
    mixPvals <- mixPvals[rearrange, rearrange]
    dimnames(mixPvals) <- list(rearrange, rearrange)
  }
  
  # combine matrices into list to return
  out <- list(Betas = round(mixCoeffs, digits = 2),
              Pvals = round(mixPvals, digits = 5))
  
  return(out)
  
}

# gather basic data (this is better as a function, then rbind)
summarizeData <- function(Data = waitData, nSubs = nSubsW, type = "Wait"){
  
  tempSummary <- data.frame(group = rep(type, nSubs))
  
  if (type == "Cognitive") {
    
    for (subj in seq(nSubs)){
      
        # Choice index
        tempChoice <- Data[[subj]]$Choice
        
        # Proportion complete
        tempSummary[subj,2] <- mean(tempChoice)
      
        # earnings
        tempSummary[subj,3] <- sum(Data[[subj]]$Offer[tempChoice==1])
        
        # RTs
        tempSummary[subj,4] <- median(Data[[subj]]$`Choice RT`[tempChoice==0])
    
        # Mistakes (only for cognitive effort group)
        tempMistakes <- sum(cogDataLong[[subj]]$`Trial Outcome`=="Forced travel") # how many forced travel trials
        tempTrialnum <- length(Data[[subj]]$Choice) # how many decision trials
        tempSummary[subj,5] <- tempMistakes / (tempMistakes + tempTrialnum) # proportion of forced travels throughout the exp        
    
    }    
    
  } else {
    
    for (subj in seq(nSubs)){
        
        # choice index
        tempChoice <- Data[[subj]]$Choice
      
        # propotion complete
        tempSummary[subj,2] <- mean(tempChoice)
        
        # earnings
        tempSummary[subj,3] <- sum(Data[[subj]]$Offer[tempChoice==1])
        
        # RTs
        tempSummary[subj,4] <- median(Data[[subj]]$RT[tempChoice==0])
        
        # Mistakes (to allow for mistakes on the cognitive group)
        tempSummary[subj,5] <- NA
        
    }
  
  }
  
  return(tempSummary)
  
}

```

``` {r Load data for between-subject experiment, echo = FALSE}
### data loading and cleaning up
# forced travels are switched to acceptances, as they reflect a preference for that trial
# rawChoice will be used to compute the # of mistakes
# The RT is upper-bounded because a glitch in the code made two 14s trials last longer (among all p's) 
setwd("./Cost2/data")
files <- dir(pattern = '_log.csv')

# load data
dataBtw <- tibble(SubjID = files) %>%
  mutate(contents = map(SubjID, ~ suppressWarnings(read_csv(., col_types = cols()))))  %>%
  mutate(Cost = substring(SubjID, 5, 8),
         Cost = case_when(Cost == "wait" ~ "Wait",
                          Cost == "cogT" ~ "Cognitive",
                          Cost == "phys" ~ "Physical",
                          Cost == "phea" ~ "Easy"),
         SubjID = as.integer(substring(SubjID, 0, 3))) %>%
  unnest() %>%
  rename(TrialN = X1) %>%
  mutate(rawChoice = Choice, 
         RT = ifelse(RT > 14.1, 14, RT),
         Choice = ifelse(Choice == 2, 1, Choice), # forced travels (2) become acceptances (1)
         Half = ifelse(Block < 4, "Half_1", "Half_2"),
         Cost = factor(Cost, levels = c("Physical", "Cognitive", "Wait", "Easy")),
         optimal = case_when(
          (Handling == 10 & Offer < 8) ~ 0,
          (Handling == 14 & Offer < 20) ~ 0,
          TRUE ~ 1
          )
         ) 

# get a simple subject list and the number of subjects
subjList_btw <- unique(dataBtw$SubjID)
nSubjs_btw <- length(subjList_btw)

```

```{r Load data for within-subject experiment, echo = FALSE}
# First looks at the new data
# The RT is upper-bounded because a glitch in the code made one 10s last 14s
setwd('./Cost3/data/')
files <- dir(pattern = 'main_log.csv')

# load the data and remove extreme subjects
dataWth <- data_frame(SubjID = files) %>% 
  mutate(contents = map(SubjID, ~ suppressWarnings(read_csv(., col_types = cols()))))  %>%
  mutate(SubjID = substring(SubjID, 0, 3)) %>%
  unnest() %>%
  mutate(RT = ifelse(RT > 10.1, 10, RT),
         Half = ifelse(Block < 4, "Half_1", "Half_2"),
         Btype = BlockType) %>%
  unite(Cost, Cost, BlockType) %>%
  rename(TrialN = X1) %>%
  mutate(rawChoice = Choice,
         Choice = ifelse(Choice == 2, 1, Choice),
         Cost = case_when(Cost == "WAIT_0" ~ "Wait_C",
                           Cost == "COGNITIVE_0" ~ "Cognitive",
                           Cost == "GRIP_1" ~ "Physical",
                           Cost == "WAIT_1" ~ "Wait_P")) %>%
  group_by(SubjID) %>%
  mutate(BlockOrder = ifelse(Btype[1] == 0, "Cognitive1st", "Physical1st"))

# Get just the subject list and number of subjects
subjList_wth <- unique(dataWth$SubjID)
nSubjs_wth <- length(subjList_wth)
```

1.  Department of Psychological and Brain Sciences, Boston University, Boston, USA

2.  Center for Systems Neuroscience, Boston University, Boston, USA

\bigskip

Corresponding authors: Claudio Toro-Serey (ctoro@bu.edu) & Joseph T. McGuire (jtmcg@bu.edu)


\newpage

## Abstract

Cognitive and physical effort are typically regarded as costly, but recent findings have suggested that exerting effort can boost the value of prospects under certain conditions. Here we embedded mental and physical effort in a "diet choice" foraging task, which required decision makers not only to evaluate the magnitude and delay of a focal prospective reward, but also to estimate the general opportunity cost of time. In two experiments, independent sets of participants collected rewards that required equivalent periods of cognitive effort, physical effort, or unfilled delay. Monetary offers varied per trial, and the two experiments differed in whether the type of effort or delay cost was the same on every trial (between-subjects, n=21 per condition), or varied across trials (within-subjects, n=48). Participants were free either to accept or reject the cost/reward prospect offered on each trial. All participants were more likely to accept offers when rewards were higher and delays shorter, in line with a reward-maximizing strategy. Participants almost never reversed their acceptance decisions, and error rates in completing the effort requirement were low and decreased over time. When participants faced only one type of cost, cognitive effort persistently produced the highest acceptance rate compared to trials with an equivalent period of either physical effort or unfilled delay. This could be because cognitive effort was directly rewarding, or because it modulated foraging-related factors such as the perceived duration of delays or the estimated richness of the environment. We theorized that if cognitive effort were intrinsically rewarding, we would observe the same pattern of preferences when participants foraged for varying cost types in addition to rewards. In the within-subject experiment, an initially higher acceptance rate for cognitive effort trials disappeared over time amid an overall decline in acceptance rates as participants gained experience with all three conditions. Our results extend the view that cognitive demands may reduce the discounting effect of delays, but also suggest that differences in cost can eventually fade if individuals actively experience alternative forms of demand. Rather than assigning intrinsic value to cognitive effort, our findings support the idea that a cognitive effort requirement might influence contextual factors such as subjective delay durations or the perceived opportunity cost of time. Such altered estimations can be recalibrated if multiple forms of demand are interleaved.


## Introduction


## Methods


## Results



## Discussion